Selecting and ranking interesting pairs of cyclic temporal granularities

### 2021-01-20 paper/main.Rmd
removed behavior of raw wpd

# Distance measure

We are interested in assessing structure in probability distributions of the measured variable across bivariate cyclic granularities. We propose a measure called Median Maximum Pairwise Distances (MMPD) to evaluate structure in such a design.

@@

# Behavior of raw wpd (weighted pairwise distances)

Most of the behavior of the measure wpd was studied via simulation. The simulations explore how wpd performs under various designs and parameters and its limitations. To study the behavior of wpd, simulations were carried out for four different designs and the following factors that could potentially have an impact on the values of wpd:

<!-- Numerous simulations were carried out in a wide range of scenarios. -->

<!-- Same scenarios were used, when possible with normalized wpd, keeping all other simulation variables constant. -->

- $nx$ (number of levels of x-axis)
- $nfacet$ (number of levels of facets)
- $\lambda$ (tuning parameter)
- $\omega$ (increment in each panel design)
- $dist$ (normal/non-normal distributions with different location and scale)
- $n$ (sample size for each combination of categories)
- $nsim$ (number of simulations)
- $nperm$ (number of permutations of data)  
- $designs$  
$D_{null}$  (No difference in distribution)  
$D_{var_f}$ (Difference in distribution only across facets)  
$D_{var_x}$ (Difference in distribution only across x-axis)  
$D_{var_{all}}$ (Difference in distribution in both facets and x-axis)

```{r simtable, eval = FALSE}
variable = c("nx", "nfacet", "lambda", "omega", "dist", "n", "nsim", "nperm", "design")

description = c("number of levels of x-axis", "number of levels of facets",
                "tuning parameter","increment in each panel design", "normal/non-normal distributions with different location and scale", "sample size", "number of simulations", "number of permutations of data", 
                "$D_{null}$ (No difference in distribution), $D_{var_f}$ (Difference in distribution only across facets), $D_{var_x}$ (Difference in distribution only across x-axis), $D_{var_all}$ (Difference in distribution in both facets and x-axis)")
table = tibble::tibble(variable, description)

table %>% 
knitr::kable(format = "latex",
             escape = TRUE) %>% 
  kableExtra::kable_styling(
    bootstrap_options = "striped", 
    full_width = F, 
    font_size = 7)

## Null design

This section explores the behavior of wpd in designs where there is no difference in distribution between x and facet categories. We have considered different initial distributions to study the impact of initial distribution under the null setup. Since the measure wpd is essentially set up to detect differences in distributions irrespective of underlying distribution, it would be ideal if it has minimal dependency on the type, location and scale of the initial distribution. To that end, some pre-processing of the data is preferred to bring it to the same scale, location and type. The Normal Score Transform or NQT has been applied in various fields of geo-statistics in order to make most asymmetrical distributed real world measured variables more
treatable and normal-like.

Following the work of Krzysztofowicz (1997) the empirical NQT involves the following steps:
1. Sort the sample of measured variable $X$ from the smallest to the largest observation $x_{(1)},\dots, x_{(i)}, .., x_{(n)}$.
2. Estimate the cumulative probabilities $p_{(1)},\dots, p_{(i)}, .., p_{(n)}$.
using a plotting position like $i/(n + 1)$ such that $p_{(i)} = P(X \leq x_{(i)})$.
3. Transforming each observation $x_(i)$ of $X$ into observation $y(i) = Q^{-1}(p(i))$ of the standard normal variate $Y$ , with $Q$ denoting the standard normal distribution and
$Q^{-1}$ its inverse, applying discrete mapping.

Further, we want to study the distribution of wpd for different $nx$ and $nfacet$.

### Simulation setup

<!-- Study the behavior of wpd under different initial distributions for different $nx$ and $nfacet$ -->

Using two types of distributions, viz. normal and gamma (non-normal), we generated observations for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{2, 3, 5, 7, 14, 20, 31, 50\}$ and $nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high. Each combination is being referred to as a _panel_. That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. That is, if we consider the panel  $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$. The values of $\lambda$ is set to $0.67$, since we want to up-weigh the within-facet distances and that of $\omega$ is set to $0$, since there is no significant differences between distributions in the null case. Observations were generated for each type of distribution changing the shape and scale to study the effect of shape, scale and type of distribution on wpd. The set of distributions considered for this purpose is ${N(0,1), N(5, 1), N(0,5), \Gamma(0.5, 1), \Gamma(2, 1)}$. Each of the scenario is run $nsim = 200$ times to see the distribution of wpd values for each scenario.

<!-- wpd is computed for different normal and non-normal distributions to see how the underlying distribution has a role to play in the value of wpd. Since we are concerned with detecting the differences in distributions, it would be ideal if wpd has minimal dependency on the type, location and scale of the  underlying distribution. To that end, some pre-processing of the data is preferred to bring it to the same scale, location and type. -->

<!-- Null design only where all combinations are equal for fixed values of $\mu$ and $\sigma$ -->
<!-- Dist 1: $N(0,1)$ individual graph -->
<!-- Dist 2: $N(\mu, 1)$ (faceted with different values of $\mu$: not yet just for one value now) -->
<!-- Dist 3: $N(0, \sigma)$ (faceted with different values of $\sigma$: not yet just for one value now) -->
<!-- Dist 4: $N(\mu, \sigma)$ faceted with different values of mu and sigma, for all levels of x and facet to see how value of wpd changes with that of change in parameters. -->

<!-- **Assumptions:** There is no difference in distribution between any facet or x-category -->
<!-- nsim = 200 -->
<!-- lambda = 0.67 -->

<!-- **Questions:**  -->
<!-- - How raw value of wpd changes with different nx and nfacet for different location and scale of a Normal and non-normal distribution -->

Figure \ref{fig:normal-diff-quantrans} and \ref{fig:gamma-diff-quantrans} show distribution of wpd for different initial normal and gamma distribution to study the effect of changing location and scale on the distribution of wpd. It seems like the  distribution changes across different facet and x levels but look the same for each panel, which implies wpd value is unaffected by the change in location and scale of the the normal distribution. Figure \ref{fig:mean-sd-ns-nfacet} shows how mean and sd of the distribution of wpd changes with the increasing x and facet levels. It seems like both mean and standard deviation are affected more by change in the x-axis levels than the facet levels. Figure \ref{fig:nxbyfacet-ridge-N05} gives another way to look at the effect of changing facet and x levels on the distribution of wpd for an initial distribution $N(0,5)$. Clearly, the location of the distribution shifts to the right for increasing x levels and scale is different for low and high values of facet levels with a longer left tail for lower facet levels. We have considered alternate distribution, location and scale for this simulation setting, but do not present the results as they are behaviorally similar.

<!-- # ```{r normal-diff, fig.cap="Ridge plots of raw wpd is shown for N(0,1) and N(5,1) distribution. The densities change across different facet and x levels but look same for the two distributions, which implies wpd value is unaffected by the change in mean and standard deviation of the normal distribution", eval = FALSE} -->
<!-- # knitr::include_graphics(here::here("simulations/null_design/figs/normal_ridge_nxbynfacet_wpd_N01.png")) -->
<!-- # ``` -->

```{r normal-diff-quantrans, fig.cap = "Ridge plots of raw wpd is shown for N(0,1), N(5,1) and N(0,5) distribution. The densities change across different facet and x levels but look same for each panel, which implies wpd value is unaffected by the change in mean and standard deviation of the normal distribution"}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_normal.png"))
```

```{r gamma-diff-quantrans, fig.cap="Ridge plots of raw wpd is shown for Gamma(0.5,1), Gamma(2,1) distribution. The densities change across different facet and x levels but look same for the two distributions, which implies wpd value is unaffected by the change in the shape paramter of the gamma distribution"}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_gamma.png"))
```

```{r mean-sd-ns-nfacet, fig.cap="Movement of mean and sd for raw wpd is shown for different number of levels (nlevel) of x-axis and facets through line plots. Mean increases and sd decreases more sharply for increasing x-axis levels as compared to facet levels. It seems like both mean and standard deviation are affected more by change in the x-axis levels than the facet levels.", fig.pos="h"}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/mean_sd_nx_nfacet.png"))

```{r nxbyfacet-ridge-N05, fig.cap="Ridge plots of raw wpd is shown for N(0,5) distribution. For each panel, it could be seen that the location shifts to the right for increasing x levels. Across each panel, the scale of the distribution seems to change for low/moderately lower values and higher values of nfacets and left tails are longer for lower facet levels."}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/nxbyfacet_ridge_wpd_N05.png"))
```

```{r gamma-diff, fig.cap="Ridge plots of raw wpd is shown for G(0.5,1) and G(2,1) distribution. The densities change across different facet and x levels and also for the two distributions, which implies wpd value is affected by the change in location value when distributions are not normal.", eval = FALSE}
knitr::include_graphics(here::here("simulations/null_design/figs/gamma_ridge_nxbynfacet.png"))
```

```{r gamma-diff-trans, fig.cap="Ridge plots of raw wpd is shown for G(0.5,1) and G(2,1) distribution after quantile transformation looks similar and hence is unaffected by change in location.", fig.pos="h"}
knitr::include_graphics(here::here("simulations/raw/null_design_quantrans/figs/diff_mean3_gamma.png"))
```

<!-- ### Choice of alternate designs -->

<!-- A normal distribution is taken for this purpose. We would calibrate the values of $WPD$ for different designs illustrated in Section \ref{sec:idea}. -->

<!-- $$\mu_{j.} = \mu + j\omega$$ -->
<!-- or, $$\mu_{.k} = \mu + k\omega$$ where,    -->
<!-- $j$ and $k$ are the index of the x and facet category respectively  -->
<!-- $\mu$ is the mean of the first category and $\omega$ is the increment from the starting value. -->

<!-- mean = 0 -->
<!-- range_mean = c(0, 1, 3,  5, 7, 15, 20) -->
<!-- sd = 1 -->
<!-- range_sd = c(1, 3,  5, 7, 15, 20) -->
<!-- range_w =  c(1, 2, 3, 4, 5) -->
<!-- range_lambda = seq(0.1, 0.9, 0.05) -->

## Alternate designs

This section explores the behavior of wpd in designs where there is infact difference in distribution between facet categories ($D_{var_f}$) or across x-categories ($D_{var_x}$) or both ($D_{var_{all}}$). Since it is established in the last section that after preprocessing the data through NQT, the initial distribution does not have a role to play in the values of wpd, we proceed with a N(0,1) distribution only. Supposably, the tuning parameter ($\lambda$) and increment parameter ($\omega$) should impact the values of wpd.

### Simulation setup

Using $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$, observations are drawn from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet =  \{2, 3, 5, 7, 14, 20, 31, 50\}$. $ntimes = 500$ is assumed for this setup as well. Furthermore, to generate different distributions across different combination of facet and x levels, the following method is deployed - suppose the distribution of the combination of first levels of x and facet category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_{var_x}$) and $\mu_{.k} = \mu + k\omega$ (for design $D_{var_f}$). Table \ref{tab:explain-design} shows an example of how initial distributions are assigned in a panel with $nfacet = 3$ and $nx = 2$ for different designs using $\omega = 1$.

```{r explain-design}
sim_varx_normal = function(nx, nfacet, mean, sd, w)
{
  rep(dist_normal((mean + seq(0, nx-1, by  = 1)*w), sd), nfacet)
}

sim_varf_normal = function(nx, nfacet, mean, sd, w)
{
  rep(dist_normal((mean + seq(0, nfacet-1, by  = 1)*w), sd), each = nx)
}
  sim_varall_normal = function(nx, nfacet, mean, sd, w)
  {
    dist_normal((mean + seq(0, (nx*nfacet - 1), by  = 1)*w), sd)
  }

mean = 0
sd = 1
w = 1
nx = 2
nfacet = 3

raw_levels = expand.grid(facet = c("$b_1$", "$b_2$", "$b_3$"),
                         x = c("$a_1$", "$a_2$"))

facet = rep(c("$b_1$", "$b_2$", "$b_3$"), each = nx)
x =  rep(c("$a_1$", "$a_2$"), nfacet)

vary_f <-  sim_varf_normal(nx, nfacet, mean, sd, w) %>% tibble()
vary_x <-  sim_varx_normal(nx, nfacet, mean, sd, w) %>% tibble()
vary_all <-  sim_varall_normal(nx, nfacet, mean, sd, w) %>% tibble()

raw_table <- bind_cols(x = x, 
                       facet = facet, 
                       vary_f = vary_f, 
                       vary_x = vary_x, 
                       vary_all = vary_all) %>% 
  tibble() 
names(raw_table) = c("x", "facet", "$D_{var_f}$", "$D_{var_x}$", "$D_{var_{all}}$" )
raw_table %>% 
  kable(format = "latex",
    booktabs = TRUE, 
    escape = FALSE,
    caption = "Simulation setup for a panel with 3 facet levels and 2 x-axis levels for different designs starting from an initial distribution N(0, 1) for the combination $(a_1, b_1)$.") %>% kable_styling()
```

<!--  - Exploring the effect of tuning parameter $\lambda$: -->

We now discuss the effect of different tuning parameters $\lambda$ and $\omega$ with the help of two alternate designs ($D_{var_x}$ and $D_{var_f}$). Figure \ref{fig:tuning-one-omega} displays different values of $\lambda$ for a relatively small and higher $\omega$ for $nx = nfacet =  \{2, 3, 5, 7, 9\}$. The $\lambda$ for which the two designs intersect will then be chosen as the optimal $\lambda$ that could then be weighed appropriately to up-weigh the within-facet distances and down-weigh the between-facet distances. The design $D_{var_x}$ increases linearly with increasing $\lambda$, whereas $D_{var_f}$ decreases linearly with increasing $\lambda$. This is expected as wpd has a linear relationship with $\lambda$ by construction. The two designs mostly intersect at $\lambda = 0.5$ for a higher value of $\omega$.  \autoref{fig:tuning-parameter-intersection} shows how the value of $\lambda$ changes with increasing $\omega$ for $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$.

<!-- Assumptions: The initial combination has a $N(\mu,\sigma)$ distribution and then every combination increases as follows: -->
<!-- - $$\mu_{j.} = \mu + j\omega$$ (for design vary_x) -->
<!-- - $$\mu_{.k} = \mu + k\omega$$ (for design vary_f) -->
<!-- Value of the parameters chosen for this simulation are $\mu = 0$, $\sigma = 1$ and $\omega = {1, 2, \dots, 10}$ -->

```{r tuning-one-omega, fig.caption = "WPD from two designs are plotted for different values of lambda for different x and facet categories. For omega = 1, the designs intersect for 0.6<lamba<=0.75, whereas for higher omega, design intersects at lambda = 0.5"}
knitr::include_graphics(here::here("simulations/tuning_param/figs/fixed_omega.png"))
```

```{r tuning-parameter-intersection, fig.cap="For most panels it is observed that the most common value of the tuning paramter for which the designs interact is 0.5, which implies any value greater than 0.5 could be chosen to up-weigh the within-facet distances and down-weigh the between-facet distances for most sitatuions."}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

How value of lambda changes with increasing omega and mean for fixed nx, nfacet and standard deviation?

```{r lambda-omega-mean, eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_mean_3_2.png"))
```

How value of lambda changes with increasing omega and sd for fixed nx, nfacet and mean?

```{r lambda-omega-sd, eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_sd_3_2.png"))
```

```{r lamba-omega-alternate-design}
knitr::include_graphics((here::here("simulations/result_report/wpd_alterate_designs.png")))
```

 Distribution across facet and x categories
<!-- ### Need for normalisation -->

From Figure \ref{fig:dist-across-facets} and \ref{fig:dist-across-x} shows the MMPD distribution is different for different levels of facets and x-axis levels. With increasing number of facets, the location of the distribution shift rightwards and with increasing x-axis levels, the scale of the distribution reduces.

Median and maximum distances are affected by the number of categories considered and hence the distance measure $MMPD_{raw}$, which is a combination of median and maximum would also be influenced by the number of levels. It would have higher values if $C_i$ or $C_j$ has higher levels. We would ideally want a higher value of the measure if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. In Figures \ref{fig:dist-across-facets} and \ref{fig:dist-across-x}, both the cyclic granularities $C_i$ and $C_j$ are considered such that their levels vary in the range $(2, 5,7 ,9, 14, 21, 30, 45)$. Each of these combinations is considered a panel for each of which $MMPD_{raw}$ has been constructed $100$ times to compute the distribution. <!--More specifically, Figure \ref{fig:mean-sd-max} shows the marginal contribution of increasing levels of $C_i$ and $C_j$ on distribution of $MMPD_raw$. This shows that with the increase in the number of levels of $C_i$, the location of the distribution shift rightward, whereas for increasing x-axis levels, the scale of the distribution decreases.-->

Therefore, in order to compare $MMPD_{raw}$ across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure.

```{r hist-nx-nfacet, eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/hist_nxbyfacet.png"))
```

```{r dist-across-facets, fig.cap = "Distribution of $MMPD_{raw}$ for different facet levels are shown. The location of the distribution shifts rightwards implying that as the number of facets increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained.", eval = FALSE}
here::here()
knitr::include_graphics(here::here("simulations/result_report","nx_by_facet_raw.png"))
```

<!-- # ```{r dist-across-facets, fig.cap = "Distribution of $MMPD_{raw}$ for different facet levels are shown. The location of the distribution shifts rightwards implying that as the number of facets increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained. "} -->
<!-- # mmpd_raw = read_rds("../hakear-drake/data/data_all23457914n.rds")  -->
<!-- # mmpd_xmore = read_rds("../hakear-drake/data/data_all213045x.rds")  -->
<!-- # mmpd_fmore = read_rds("../hakear-drake/data/data_all_213045f_allx.rds") -->
<!-- #  -->
<!-- # data_all = bind_rows(mmpd_raw, mmpd_xmore) -->
<!-- #  -->
<!-- # data_long = data_all %>% pivot_longer(c(1:2), names_to = "category", values_to = "levels") -->
<!-- #  -->
<!-- # g1 <- data_all %>%  -->
<!-- #   filter(!(nx %in% c(3, 4))) %>%  -->
<!-- #   ggplot() + -->
<!-- #   ggridges::geom_density_ridges(aes(x = raw_mmpd,  -->
<!-- #                                     y = as.factor(nfacet))) + -->
<!-- #   facet_wrap(~nx, nrow = 2) -->
<!-- # g1 -->
<!-- # ``` -->

```{r dist-across-x, fig.cap = "Distribution of $MMPD_{raw}$ for different x levels are shown. The scale of the distribution becomes less and less implying that as the number of x levels increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained as the spread of the distribution decreases. ", eval = FALSE}
knitr::include_graphics(here::here("simulations/result_report/","nfacet_by_nx_raw.png"))
```

<!-- # ```{r dist-across-x, fig.cap = "Distribution of $MMPD_{raw}$ for different x levels are shown. The scale of the distribution becomes less and less implying that as the number of x levels increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained as the spread of the distribution decreases. "} -->
<!-- # g2 <- data_all %>%  -->
<!-- #   filter(nfacet<=14, nfacet!=3) %>%  -->
<!-- #   ggplot() + -->
<!-- #   geom_boxplot(aes(y = raw_mmpd,  -->
<!-- #                                     x = as.factor(nx))) + -->
<!-- #   facet_wrap(~nfacet) + coord_flip() -->
<!-- #  -->
<!-- # #  hist_plot = data_all %>%  -->
<!-- # #       ggplot() + -->
<!-- # #       geom_density(aes(x = raw_mmpd), fill = "blue") + -->
<!-- # #       facet_grid(nx~nfacet) -->
<!-- # # hist_plot -->
<!-- # g2 -->
<!-- # ``` -->

<!-- both mean and standard deviation of the distribution of maximum and median varies for different sample sizes ($n$). There is a rightward shift in the location and decrease in scale with increasing $n$. -->

## Sample size
## Number of permutations

# Normalisation

In an attempt to make the densities \ref{fig:hist_plot} same across different levels, we want to start by making their scale and location same. In that regard, we shuffle the data multiple times and compute the sampling distribution of $mmpd_{raw}$ for each combination of facet and x-axis levels.

The null hypothesis is that the two cyclic granularities do not differ on the outcome (i.e., that the outcome is observed independently of treatment assignment). When we permute the outcome values during the test, we therefore see all of the possible alternative treatment assignments we could have had and where the mean-difference in our observed data falls relative to all of the differences we could have seen if the outcome was independent of treatment assignment. While a permutation test requires that we see all possible permutations of the data (which can become quite large), we can easily conduct “approximate permutation tests” by simply conducting a vary large number of resamples. That process should, in expectation, approximate the permutation distribution.

Finally, we run some simulation experiments to see if normalisation works. If normalisation has worked, it should work for both x levle and facet-levels.

Thus these median maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels, consequently enabling comparison across different harmonies. The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.

```{r, eval = FALSE}
knitr::include_graphics(here::here("simulations/null_design/figs/normalised","gamma_ridge_nxbynfacet.png"))
```

The mean and sd for each combination of facet and x levels could be computed by shuffling the data repeatedly and obtaining the distribution of $MMPD_{raw}$ for different combinations. If $MMPD_{raw}$ is then adjusted by the location and scale of this distribution, we could assume that the distribution of the resultant $MMPD_{norm}$ will have same distribution across different combinations of x and facet levels. All data is repeatedly shuffled $nperm$ times in random manner to obtain the measure $MMPD_{raw}$. Then the $MMPD_{norm}$ is obtained by scaling the observed value by the mean and sd of the distribution of $MMPD_{raw}$ obtained from the permuted data.

\noindent Step 7 in the Algorithm defined in Section \ref{sec:} need to be revised in order to compute $MMPD_{norm}$ instead of $MMPD_{raw}$. Step 1-6 stays the same. The entire algorithm will have to repeated for all harmony pairs considered in the context.

A simulation study is conducted to see if this methodology is working.

<!-- #### Method 2: -->

<!-- $m' = max_{x}dist$ is obtained for each facet. -->
<!-- All $x$ levels are shuffled for each facet to obtain several $m'$s. Hence $m'$ is normalized for each facet level and the facets are aggregated using $m''= median(m')$, which is basically median(norm(max)). We need to normalise this median across facets as well. So we shuffle all facets for each $x$ level and obtain several $m''$s. In this approach, data is shuffled $2*nperm$ times and hence distances are computed $2*nperm$ times. It is computationally very intensive. -->

<!-- Hence, it was decided that the first approach will be employed to implement normalisation. One of the advantages of the second method could be to have normalisation separately for max and median measure. There is nothing that is stopping the users to implement this with more efficient computing power.  -->

<!-- The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the harmonies. -->

<!-- With increasing n, mean and sd of distribution of max increases. -->

<!-- FTP theorem works for large n -->

<!-- For small n resort to bootstrapping -->

<!-- bullet points end here -->

<!-- Suppose that $X_1$, $X_2$ , ... , $X_n$ are i.i.d. random variables with expected values $E(X_i) = \mu < \infty$ and variance $Var(X_i) = \sigma^2 < \infty$. Let  -->
<!-- $Y = max(X_1, X_2, \dots, X_n)$. -->

<!-- Let $F_X(x)$ be the common distribution of the variables $X_i$ and let $F_Y(y)$ be the corresponding distribution of $Y$. $F_Y(y)$ could be obtained from $F_X(x)$ simply by using: -->
<!-- $F_Y(y) = P[(X_1 \leq y)\cap(X_2 \leq y) \cap...\cap (X_n \leq y)] = {F_X(y)}^n$. For large $n$, the distribution of $Y$ approaches a standard shape, which does not depend on $F_X$. But what about the case when $n$ is not large enough? The distribution of maximum in that case will indeed depend on $n$ and the underlying distribution of $X$. If $F_X(x)$ is the CDF of $X$, then $F_Y(y) = {F_x(y)}^n$. Suppose $\Phi$ nd $\phi$ are the cdf and pdf of a standard normal distribution, then -->
<!-- $f_Y(y) = n{\Phi(y)}^{n-1}\phi(y)$, which depends on $n$. Hence, we are trying to normalise for $n$. Also, it depends on the underlying distribution of $X$, which we have assumed as normal in our case. As $n$ grows, we  can see the right tail growing, which implies that the probability that we will get a higher maximum is more. Now, for large $n$, we used EVT to normalise for $n$, that is, we brought them to the same scale without distorting the range of the distribution. But in our case, we will mostly have small $n$. It is important to ensure that they have the same mean and variation, for being able to compare the maximum value across $n$. We observe from the following graphs that our normalisation works after $n=6$, after which the difference in mean and standard deviation flattens out a lot. -->

<!-- ### why normalize -->

<!-- Median and maximum distances are not robust to the number of levels and the metric would have higher values for harmonies with higher levels. Figure \ref{fig:mean-sd-max} shows how both mean and standard deviation of the distribution of maximum and median varies for different sample sizes ($n$). There is a rightward shift in the location and decrease in scale with increasing $n$. Thus these median maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels, consequently enabling comparison across different harmonies. The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process. -->

<!-- More formally, $d_{1},d_{2}\ldots ,d_{n}$ be a sequence of independent and identically-distributed pairwise distances and $M_{n}=\max\{d_{1},\ldots ,d_{n}\}$. Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{n}, b_{n})$ exists such that each $a_{n}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family. The normalizing constants $(a_{n}, b_{n})$ vary depending on the underlying distribution of the pairwise distances. Hence to normalize appropriately, it is  -->
<!-- important to assume a distribution of these distances.  -->

<!-- For large $n$, the distribution of $M$ approaches a standard shape. But what about the case when $n$ is not large enough? The distribution of maximum in that case will indeed depend on $n$ and the underlying distribution of $D$. For large $n$, we could use EVT and CLT to normalise for $n$, however, most of the cyclic granularities encountered in practice have low $n$. Hence, normalisation could be done by bootstrapping procedure to ensure that the distribution of maximum/median have the same mean and variation for varying $n$. -->

<!-- ```{r create_simdata} -->

<!-- ```{r create_nlevels} -->

<!-- ```{r norm_max_new} -->

<!-- ```{r mean-sd-max, fig.cap = "Display shows mean and standard deviation of the distribution of maximum for sample sizes ranging from 2 to 30. There is a right shift in the distribution of maximum and median with increasing sample sizes."} -->

<!-- library(ggridges) -->
<!-- library(viridis) -->
<!-- max_all_data <- create_nlevels(nlevels = seq(2, 30, 2), -->
<!--                            nsim = 500, -->
<!--                            sim_dist = distributional::dist_normal(5, 10), create_fun = max) -->

<!-- g1 <- max_all_data %>% ggplot() + geom_density_ridges(aes(x=sim_data, y = as.factor(ind))) + ylab("n") + xlab("max") -->

<!-- median_all_data <- create_nlevels(nlevels = seq(2, 30, 2), -->
<!--                            nsim = 500, -->
<!--                            sim_dist = distributional::dist_normal(5, 10), create_fun = median) -->

<!-- g2 <- max_all_data %>% ggplot() + geom_boxplot(aes(x=sim_data, y = as.factor(ind)))  + ylab("n") + xlab("median") -->

<!-- ggpubr::ggarrange(g1, g2, ncol = 2) -->

@@

<!-- ### Distribution of distances -->

<!-- ### Theoretical evidence -->

<!-- JS distances are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$. -->
<!-- With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made. -->

<!-- ### Empirical evidence -->

<!-- Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution. We look at different scenarios, where observations are collected from Normal, Exponential, Chi-squared and Gumbel distribution and found the distribution of JS distances are similar, irrespective of which distribution they are drawn from. -->

<!-- #### Initial distribution of observed variables shown in plot title -->

<!-- ```{r distv11} -->

<!-- ```{r distv12} -->

<!-- ```{r distv13} -->

<!-- ```{r distv14} -->

<!-- ### Design and analysis of simulation experiments -->

The behavior of the measure is monitored and understood through simulation experiments. A single simulation consisting of computational operations on a panel generating MMPD the value of which represents to what extent a pair of cyclic granularities would be interesting when displayed in the design. In comparing different simulation runs, we distinguish between different distributions and simulation scenarios (also iterations may be).

### Simulated designs

 D1. Same distribution of the measured variable across all x-axis and facet categories
 D2. Different distribution across facet categories but same across x-axis categories
 D3. Different distribution across x-axis categories but same across facet categories
 D4. Different across both x-axis and facet categories

Each of these case scenarios tried against Normal and a non-normal (Gamma) distribution to check if underlying distribution has a role to play.

### Environment

R version 4.0.1 (2020-06-06) is used with platform: x86_64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6

### Experimental set up

This section describes the design of simulation runs, in terms of the scenarios simulated, the number of permutations used to compute estimates and to display the distribution of MMPD.

 <!-- - seed while generating random observations `seed1 = set.seed(9999)`   -->
 <!-- - grid with how many x levels `range_nx_vec = seq(2:4)`   -->
 <!-- - grid with how many facet levels `range_nfacet_vec = seq(2:4)`   -->
 <!-- - number of observations to be generated `ntimes_val = 500`   -->
 <!-- - number of permutation of data done to compute estimates `nperm_val = 2`   -->
 <!-- - number of permutations of data done to display histogram for each panel `nsim_val = 10`    -->
 <!-- - ordered or un-ordered categories of x levels while computing distances `dist_ordered_val = TRUE`   -->
 <!-- - distance between which quantiles are computed `quantile_prob_vec = seq(0.01, 0.99, 0.01)`   -->

```{r}
knitr::include_graphics(here::here("paper/Figs/simulation.png"))
```

Observations from Normal(5, 10) and Gamma(0.5, 0.2) are3 chosen in all the designs.

<!-- . the null distribution are chosen as:   -->
<!-- `sim_norm_null =   distributional::dist_normal(5, 10)`   -->
<!-- <!--`sim_exp_null  =   distributional::dist_expoential(5)`  --> -->
<!-- `sim_gamma_null  =   distributional::dist_gamma(0.5, 0.2)`   -->
<!-- <!--`sim_weibull_null  =   distributional::dist_weibull(1.5, 3)`-->   -->

<!-- The various designs are selected by increasing the mean/rate/shape of these distributions proportionally with that parameter. -->

<!-- To check if different distributions impact, simulate data from different distributions. A 3 way contingency table with number of facet levels, x-axis levels and distribution type should be constructed with each cell value as the p-value. -->

<!-- Table -->
<!-- Conclusions -->
<!-- (in Null distribution) -->

<!-- #### Varying distribution across facet -->
<!-- #### Varying distribution across x-axis -->
<!-- #### Varying distribution across both facets and x-axis -->
<!-- #### Repeat all with varying facet and x-axis levels -->

<!-- Table -->
<!-- Conclusions -->
<!-- (in power) -->

<!-- ```{r} -->
<!-- raw_data = rnorm(12, 5, 10) %>%  -->
<!--   as_tibble() %>%  -->
<!--   mutate(nfacet = rep(1:4, each = 3), -->
<!--          nx = rep(1:3, 4)) -->

<!-- raw_data %>% group_by(nfacet, nx) %>% nest() -->

<!-- raw_data2 = raw_data %>% pivot_wider(id_cols = 2, names_from = "nx", values_from = "value") -->

<!--  ### Original -->
<!-- | facets/x-axis     	| A_1  	| A_2  	| A_3  	| .. 	| .. 	| A_m  	| -->
<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
<!-- | B_1 	| x_1 	| x_2 	| x_3 	| .. 	| .. 	| x_n 	| -->
<!-- | B_2 	| y_1 	| y_2 	| y_3 	| .. 	| .. 	| y_n 	| -->
<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
<!-- | B_L 	| z_1 	| z_2 	| z_3 	| .. 	| .. 	| z_m 	| -->

<!-- ### Step1 -->

>>>>

# The distance measure

We are interested in assessing structure in probability distributions of the measured variable across bivariate cyclic granularities. We propose a measure called Median Maximum Pairwise Distances (wpd) to evaluate structure in such a design.

@@

## Normalisation

@@

# Choosing harmonies with significant wpd
